---
title: "Exercise one"
output:
  html_document:
    df_print: paged
---


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


# Pencil-and-paper problem
## Problem1
### (a)

\begin{equation}
\epsilon = (ln(\alpha/2)/-2*n)^0.5
width= 2n\epsilon
\end{equation}


```{r}
get_ci <- function(n,alpha,p){
  e=((log(alpha/2))/(-2*n))^0.5
  lower_b= n*(p-e)
  upper_b= n*(p+e)
  ci= c(lower_b,upper_b)
  ci
}
```


### (b)
```{r}
set.seed(1234)
for (n in c(10,100,1000)){
  for (p in c(0.5,0.9,0.99)){
    w <- rbinom(10000,n,p)
    ci=get_ci(n = n,alpha = 0.05,p = p)
    #print(ci)
    no <- sum(w < ci[2] & w>ci[1])
    print(no)
  }
}


```





# Computer problems
```{r}
library(tidyverse)
```

## Problem2

```{r}
library(ISLR)
names(College)
fix(College)
summary(College)
pairs(College[,1:10])
boxplot(College$Outstate~College$Private)

Elite <- College$Top10perc > 50

summary(Elite) 
boxplot(College$Outstate~Elite)
for (bin_no in c(3,5,10)){
    par(mfrow=c(2,2))
  hist(x = College$PhD,breaks = bin_no)
  hist(x = College$S.F.Ratio,breaks = bin_no)
  hist(x = College$Books,breaks = bin_no)
  hist(x = College$Grad.Rate,breaks = bin_no)

}


# I made a mistake here that I should bin the same variable together to see how dis change with different breaks
```

## Problem3
### (a)
```{r}
set.seed(1234)
xi <- runif(30,-3,3)
#hist(xi,breaks = 5)
ei <- rnorm(30,0,0.4)
#hist(ei,breaks = 5)
yi <- 2+xi-0.5*xi^2 +ei
df <- data.frame(xi,yi)

p_list <- vector(mode = "list",11)
mse_vec <- 1:11
M_list <- vector(mode = "list",11)
for (i in 1:11){
  M <- lm(yi ~ poly(x = xi,degree = i))
  mse <-  MLmetrics::MSE(predict(M),yi)
  p <- df %>% mutate(pred= predict(M)) %>% 
    ggplot()+
    geom_point(aes(xi,yi))+
    geom_line(aes(xi,pred))+
    ggtitle(paste0("MSE = ",  mse))
            
  p_list[[i]] <- p
  mse_vec[i] <- mse
  M_list[[i]] <- M

}

p_list
plot(0:10,y = mse_vec,type = "bar",xlab = "polynomial K", ylab = "training MSE")
```

### (b)

```{r}
set.seed(1234)
xi  <- runif(1000,-3,3)
#hist(xi,breaks = 5)
ei  <- rnorm(1000,0,0.4)
#hist(ei,breaks = 5)
yi <- 2+xi -0.5*xi^2 +ei
df  <- data.frame(xi,yi)

p_list <- vector(mode = "list",11)
mse_vec_test <- 1:11
for (i in 1:11) {
  pred <- predict(M_list[[i]], newdata= df )
  mse <-  MLmetrics::MSE(y_pred = pred, y_true =  yi )
  p <- df  %>% mutate(pred= pred) %>% 
    ggplot()+
    geom_point(aes(xi ,yi ))+
    geom_line(aes(xi ,pred))+
    ggtitle(label = paste0("MSE = ",  mse) )
    
  p_list[[i]] <- p
  mse_vec_test[i] <- mse

}
 p_list
 
 plot(0:10,y = mse_vec,type = "bar",xlab = "polynomial K", ylab = "test MSE")

data.frame(poly_order= 0:10, training_mse = mse_vec, test_mse= mse_vec_test) %>% 
  tbl_df %>% 
  gather(key = MSE_type,value = ERROR,2:3) %>% 
  ggplot() +
  geom_point(aes(x=poly_order,y = ERROR,color=MSE_type))+
  geom_line(aes(x=poly_order,y=ERROR,color=MSE_type))


 
```



### (c)
```{r}
# the reason 2 poly are not highest is that after CV the test error for higher p goes down and errors are some how randomly distributed so by chance it may not be the highest.
set.seed(1234)
xi <- runif(30,-3,3)
#hist(xi,breaks = 5)
ei <- rnorm(30,0,0.4)
#hist(ei,breaks = 5)
yi <- 2+xi-0.5*xi^2 +ei
idx <- rep(1:10,3)
df <- data.frame(xi,yi,idx)
mse_mat <- matrix(1:110,nrow = 10, byrow = T)
for (i in 1:11){
  for(j in 1:10){
    M <- lm(yi ~ poly(x = xi,degree = i),data = df %>% filter(idx!=j))
    mse_mat[j,i] <-  MLmetrics::MSE(y_pred = predict(M,df %>% filter(idx==j)),
                                    y_true = df %>% filter(idx==j) %>% pull(yi)
                                            )
  }
}

i=1
j=1
df %>% filter(idx!=j)

data.frame(K= as.integer(0:10),MSE= apply(mse_mat,2,sum)) %>% 
  tbl_df() %>% 
  ggplot(aes(x=K,y=MSE))+
  geom_point()+
  geom_line()

data.frame(K= as.integer(0:10),MSE= apply(mse_mat,2,sum)) %>% 
  tbl_df() %>% arrange(MSE)

seq(1,30,3)
```

